---
title: "Factor Analysis"
output:
  html_notebook: default
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: sentence
---

# Introduction

- <a href= '#understanding'>Understanding the Dataset</a>
- <a href= '#cormatrix'>Correlation Matrix</a>
  - <a href= '#corplot'>Correlation Plot</a>
- <a href= '#factor-analysis'>Factor Analysis</a>
  - <a href= '#kmo-barlett'>KMO and Bartlett's Test</a>
    - <a href= '#kmo-test'>KMO Test</a>
    - <a href= '#barlett-test'>Bartlett's Test</a>
    - <a href= '#anti-image'>Anti-image Correlation/Covariance Matrix</a>
  - <a href= '#scree-plot'>Scree Plot</a>
    - <a href= '#parallel-analysis'>Parallel Analysis Scree Plots</a>    
  - <a href= '#rotated-compmatrix'>Rotated Component Matrix</a>    
    - <a href= '#graph-factor-loading'>Graph Factor Loading Matrices</a>  
  - <a href= '#cronbach-alpha'>Cronbach's Alpha</a> 
- <a href= '#conclusion'>Conclusion</a> 
 

    
<a id='understanding'></a>

# Understanding the Dataset

```{r}
# R Packages
packages <-
  c("tidyverse",
    "ggplot2",
    "rstatix",
    "pander",
    "heplots",
    "readr",
    "psych",
    "corrplot")

suppressPackageStartupMessages(lapply(packages, require, character.only = TRUE))

# Importing Data
df <- read.csv("./dataset/data.csv", sep = ";")


# for kaggle
# df <- read.csv("../input/airline-passenger-satisfaction/train.csv, sep = ";")

# Top 6 rows
head(df)
```

```{r}
class(df)
```

```{r}
# Structure of the data frame
str(df)
```

```{r}
print(sapply(df, class))
```

```{r}
# To convert all integer columns to numeric columns

df <- df %>% mutate_if(is.integer, as.numeric)

print(sapply(df, class))
```

```{r}
dim(df)
```

```{r}
# Structure of the data frame
str(df)
```

```{r}
# checking for any missing values in a dataframe
any(is.na(df))
```

```{r}
# Count of missing values
sum(is.na(df))
```

```{r}
#Position of missing values by column
sapply(df, function(x) which(is.na(x)))

df$ArrivalDelayinMinutes[is.na(df$ArrivalDelayinMinutes)] <- 0

any(is.na(df))
```

```{r}
# Statistical summary
summary(df)
```

<a id='cormatrix'></a>

# Correlation Matrix

Correlation matrix is a table showing correlation coefficients between variables.
Correlation coefficient is a statistical measure of the strength of a linear relationship between two variables.

```{r}
data_01 <- df %>%
  as.data.frame() %>%
  dplyr::select(09:22)

data_cor <- data_01 %>%
  as.matrix() %>%
  cor() 
```

```{r}
view(list(data_cor))
```


<a id='corplot'></a>

## Correlation Plot

```{r}
corrplot(data_cor,
         method = "number",
         col = colorRampPalette(c("white", "#85C1E9", "#4A235A"))(100))
```

```{r}
corrplot(data_cor,
         method = "color")
```


<a id='factor-analysis'></a>

# Factor Analysis

Factor analysis is a way of taking a chunk of data and shrinking it into a smaller dataset that is more manageable and more understandable.


<a id='kmo-barlett'></a>

## KMO and Bartlett's Test


<a id='kmo-test'></a>

### KMO Test

The Kaiser--Meyer--Olkin (KMO) test is a test used to examine the strength of the partial correlation between variables.
So it measures how convenient it is.

```{r}
# Kaiser-Meyer-Olkin Measure of Sampling Adequacy
KMO(data_cor)
```

<u>**KMO Test Interpretation**</u><br>

If the Kaiser--Meyer--Olkin (KMO) Sampling Adequacy Measure is greater than or equal to 0.60, factor analysis is continued.
The result is 0.78, which is sufficient.

<a id='barlett-test'></a>

### Bartlett's Test

Bartlett's test of Sphericity tests the null hypothesis that the correlation matrix is an identity matrix.
That is, it is used to test the hypothesis that the correlation matrix is a unit matrix.

```{r}
# Bartlett's Test of Sphericity
cortest.bartlett(data_cor, n = nrow(df))
```

<u>**Bartlett's Test Interpretation**</u><br>

An important statistical test (usually less than 0.05) shows that the correlation matrix is not really a unit matrix (rejection of the null hypothesis).
If the result is less than 0.05, the analysis is continued.


<a id='anti-image'></a>

### Anti-image Correlation/Covariance Matrix

The image of a variable is defined as that part which is predictable by regressing each variable on all the other variables.
So anti image is the part of the variable that cannot be predicted [ref](https://www.stata.com/manuals13/mvglossary.pdf).

```{r}
# Anti-image Covariance

view(KMO(data_cor)$ImCov)
```

```{r}
# Anti-image Correlation
library(multiUS)
view(antiImage(X = data_01))
```

<a id='scree-plot'></a>

## Scree Plot

It is the plot of the eigenvalues and the number of factors in the order of subtraction.
It is used to determine the optimum number of factors to keep in the final solution.

```{r}
scree(data_cor, factors = FALSE)
```

<a id='parallel-analysis'></a>

### Parallel Analysis Scree Plots

Parallel analysis scree plot is a technique designed to help take some of the subjectivity out of interpreting.

```{r}
parallel <- fa.parallel(data_01)
```

<a id='rotated-compmatrix'></a>

## Rotated Component Matrix

The rotated component matrix helps in determining what the components represent.

```{r}
data.fa <- principal(data_cor, nfactors = 3, rotate = "varimax")

data.fa$loadings
```

```{r}

print.psych(data.fa, sort = TRUE)


#ICLUST.sort(data.fa$loadings,labels = NULL,keys = FALSE,clustsort = TRUE)

```

-   **Extraction Method:** Principal Component Analysis.

-   **Rotation Method:** Varimax with Kaiser Normalization.

    -   Rotation converged in 4 iterations.

Varimax Rotation: It is used as a statistical technique used at one level of factor analysis as an attempt to clarify the relationship between factors.
In short, it tries to redistribute factor loads such that each variable measures exactly one factor.

<u>**NOTE**</u><br> For *IBM SPSS Statistics* : `Analyze > Dimension Reduction > Factor`.
You can perform the analysis by ticking the ones you want in the "Options" section.



<a id='graph-factor-loading'></a>

### Graph Factor Loading Matrices

```{r}
fa.diagram(data.fa)
```

<u>**Interpretation**</u><br>

v6 and v12 is a low-contribution variable.



<a id='cronbach-alpha'></a>

## Cronbach's Alpha

It helps measure the internal consistency of a set of data.
It is a reliability coefficient.
Provides verification of the consistency of a questionnaire or survey.

```{r}
library(ltm)

rc1_alpha <- data_01[, c("v14",  "v7", "v5", "v8")]
rc3_alpha <- data_01[, c("v13", "v11", "v9", "v10")]
rc2_alpha <- data_01[, c("v3", "v1", "v4", "v2")]


# rc = rotated component
cronbach.alpha(rc1_alpha)
cronbach.alpha(rc3_alpha)
cronbach.alpha(rc2_alpha)
```

| alpha   |           |
|---------|-----------|
| \>=0.9  | Excellent |
| 0.8-0.9 | Good      |


<a id='conclusion'></a>

# Conclusion

Factor 1: **Comfort**

-   Cleanliness
-   Seat Comfort
-   Food And Drink
-   Inflight Entertainment

Factor 2: **Service**

-   Ease Of Online Booking
-   Inflight Wifi Service
-   Gate Location
-   Departure Arrival Time Convenient

Factor 3: **Convenience**

-   Inflight Service
-   Baggage Handling
-   On Board Service
-   Leg Room Service